FAIR Theory: Applying Open Science Principles to the Construction and Iterative Improvement of Scientific Theories

### 2024-11-20 manuscript.Rmd
update 2024-11-20 11:57:25.770558

There are clear indications that psychological theory has limited interoperability. 
For one, theories are rarely *"refuted nor corroborated, but instead merely fade away as people lose interest"* [@meehlTheoreticalRisksTabular1978].
To be interoperable, psychological theory would need to play a concrete part in scientists' day-to-day work.
For example, it should be possible to integrate theory directly into analysis workflows:
to derive hypotheses, select control variables, and guide model specification.
The aforementioned lack of formalization [REF Robinaugh] and ambiguity [REF Frankenhuis] prevent interoperability in such a practical sense.
Additionally, theories should be interoperable with each other:
for instance, it should be possible to embed a specific theory about the process of emotion regulation [REF Gross] within a theory of emotion regulation development [REF Morris].

>>>>

Interoperability pertains to the property of digital research artefacts to "integrate or work together [...] with minimal effort" [@wilkinsonFAIRGuidingPrinciples2016a].
As the original definition of interoperability was narrow and somewhat abstract,
the concept has recently been further refined in terms of facilitating "successful communication between machines and between humans and machines", where "A and B are considered X-interoperable if a common operation X exists that can be applied to both" [@vogtFAIR20Extending2024].

There are clear indications that psychological theory has limited interoperability. 
For one, theories are rarely *"refuted nor corroborated, but instead merely fade away as people lose interest"* [@meehlTheoreticalRisksTabular1978].

To be interoperable, psychological theory would need to play a concrete part in scientists' day-to-day work.

For example, it should be possible to integrate theory directly into analysis workflows:
to derive hypotheses, select control variables, and guide model specification.
The aforementioned lack of formalization [REF Robinaugh] and ambiguity [REF Frankenhuis] prevent interoperability in such a practical sense.
Additionally, theories should be interoperable with each other:
for instance, it should be possible to embed a specific theory about the process of emotion regulation [REF Gross] within a theory of emotion regulation development [REF Morris].

### Reusability
